---
title: "PUBH7475 Homework V"
author: "Renchang Lu 5504577"
date: April 21, 2019
output: 
  html_document:
    theme: cerulean
    highlight: tango
    df_print: kable
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### **Introduction**

<font size=3>
In this paper, I will apply **SVD and Funk SVD** to 100K MovieLense data set. The 100K MovieLens data consists of 1,000,000 anonymous ratings on a five-star scale from 1,000 users on 1,700 movies. 
To be specific, I will divide the data set into training set, known test set and unknown test set. Then, use "recommenderlab" library to build a recommend system with method Popular and UBCF separately for ignorable and nonignorable data. At last, evaluate the performance by RMSE.
</font> 

&nbsp;

#### **Data Processing**

<font size=3>
In this part, I load "tidyverse" package at first, which is good at data processing. Also. load "recommenderlab" package, which is for recommender system construction. Then, I load "MovieLense" data set and do 5 fold cross validation. Then, I get the training data set, known test data set and unknown test data set.
</font> 

```{r, load_packages, include = FALSE}
# Loading packages
library("tidyverse") 
library("recommenderlab")
library(reshape2)
library(ggplot2)
```

```{r}
# Load the data set
data("MovieLense")
movielense = MovieLense
# Set seed to fix the result
set.seed(101)
# k value for k fold cross validation
n_fold <- 5

# Do 5 fold cross validation
eval_sets <- evaluationScheme(data = movielense, method = "cross-validation", k = n_fold, 
                              given = -1)

eval_sets

```

```{r}
# Get training set
evaltrain <- getData(eval_sets, "train")
# Get known test set
evalknown <- getData(eval_sets, "known")
# Get unknown test set
evalunknown <- getData(eval_sets, "unknown")  
```


#### **Visualization**
<font size=3>
In this part, I did some visualization to know more about the data set.

```{r, fig.align='center'}
## visualize part of the matrix
image(MovieLense[1:200,1:200])
## number of ratings per user
hist(rowCounts(MovieLense))
## number of ratings per movie
hist(colCounts(MovieLense))

```

From the graphs, it is obvious that this data set is sparse. Large amounts of users and movies do not have rating.
</font> 

&nbsp;

#### **Creating recommender system model**

<font size=3>
In this part, I'll build a SVD recommender system at first, and calculate the RMSE.

```{r}
## Build SVD
algorithms <- list(SVD = list(name = "SVD", param = NULL))
model_svd <- Recommender(evaltrain, method = "SVD")

## Predict preference scores of each user.
svd_pre <- predict(model_svd, evalknown, type="ratings")
## Show the prediction of first 5 users rates to 5 movies. 
as(svd_pre, "matrix")[1:5,1:5]

## Calculate RMSE
ev_svd <- evaluate(eval_sets, algorithms, type="ratings")
svd_accuracy <- avg(ev_svd)

svd_accuracy
```

The RMSE of SVD recommender system is 1.0241. It seems high because I didn't consider the missing data.

```{r, fig.align='center'}
moviematrix <- as(movielense, "matrix")
## Calculate standard deviation of the rating of each user
user_std <- c()
for (i in 1:nrow(moviematrix))
  user_std[i] <- sd(as(movielense, "matrix")[i,], na.rm=TRUE)
## Draw the histogram
hist(user_std)
```
From the histogram, we can find that the standard deviation of each user's rating is approximately followed normal distribution, which proves that missing does not occur at random.

</font> 

&nbsp;

#### **Recommender system model with nonignorable missing data**

<font size=3>
In this part, I'll build Funk SVD to deal with the nonignorable missing data to see how much predictive performance can be enhanced by incorporating the missing pattern.

```{r}
## Build Funk SVD
algorithms2 <- list(SVDF = list(name = "SVDF", param = NULL))
model_SVDF <- Recommender(evaltrain, method = "SVDF")
## Predict preference scores of each user.
SVDF_pre <- predict(model_SVDF, evalknown, type="ratings")
## Show the prediction of first 5 users rates to 5 movies. 
as(SVDF_pre, "matrix")[1:5,1:5]
## Calculate RMSE
ev_SVDF <- evaluate(eval_sets, algorithms2, type="ratings")
SVDF_accuracy <- avg(ev_SVDF)

SVDF_accuracy

```

The RMSE now is 0.9339, it proves that the predictive performance can be enhanced by incorporating the missing pattern.

</font> 



```{r}
moviematrix <- as(movielense, "matrix")
user_std <- c()
for (i in 1:nrow(moviematrix))
  user_std[i] <- sd(as(movielense, "matrix")[i,], na.rm=TRUE)
```





